Further Explorations of \suldiox~Aqueous Adsorption:

### 2011-09-08 Background.tex
More of Fred's final edits

In this work we adopt Buch et al's nomenclature to \suldiox~in order to quantify hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions from the \suldiox-sulfur to water-oxygens. Thus, an ``SOO'' coordinated \suldiox~molecule forms a single interaction through the sulfur atom to a neighboring water-oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox-oxygens. Analysis of the distribution of \suldiox~coordinations will give insight to how \suldiox~binds to the water surface.

In our analysis we defined bonds using the distance criteria of Baer et al. in determining the \suldiox~bonding coordinations.\cite{Baer2010} The bond-length definition is based on a set of distance criteria where a bonding interaction between a \wat-oxygen and \suldiox-sulfur is formed at a distance less than 3.5 \angs, and an \suldiox-oxygen hydrogen bond to a \wat-hydrogen is formed at a distance less than 2.2 \angs.

@@

Hydrated \suldiox~clusters have been studied extensively with many recent experiments and computations forming a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayashi1985, Moin2011, Eckl2008} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The first attempt to elucidate the structure of surface hydrated \suldiox~by Baer et al. showed that the bonding coordination distributions of \suldiox~are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''. In the same work they then focused on the most dominant coordination to determine the most likely cluster geometry using two and three-water clusters bound to the \suldiox.

The cluster geometries found in this study imply a cyclic bonding structure through the two or three waters involved. ``Cyclic'' here is used to denote a closed loop formed by the intermolecular hydrogen bonds, S-O interactions, and covalent bonds of the molecules involved. Figure \ref{fig:cyclic-example} depicts one such cyclic structure showing the bonds beginning on the sulfur and returning through the \suldiox-oxygen.

@@

\subsection {Graph Theoretical Details}
The optimized geometry of the \suldiox-hydrates suggests cyclic bonding structures, but it remains a different story entirely when \suldiox~is placed in a dynamic environment such as in the course of molecular dynamics simulations of an aqueous surface. Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters in a static vacuum. Do the cyclic structures also form in the course of a dynamic bonding process on a simulated water surface, where extended hydrating structures influence \suldiox~and water behavior? To study the formation and behavior of cyclic hydrate structures we employ graph theoretical techniques on MD trajectory data. Previous use of graphs in molecular computations were applied to finding stable arrangements of water clusters, ice, hydrogen bonding, extracting topological molecular properties, and cyclic structure studies.\cite{Anick2002, Huber2007, Radhakrishnan1991, Shi2005, Garcia2004, McDonald1998}

@@

To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A BFS is a recursive algorithm of queuing nodes and all neighboring nodes while performing a specified procedure on each visited node. This is easily performed on adjacency list or connectivity matrix data structures, iterating through nodes (i.e. atoms) of interest in the graph as starting points of the search. In BFS terminology, all nodes are colored during graph traversal to distinguish unvisited nodes (white), queued nodes (gray), and visited nodes (black). Using a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queuing adjacent neighbors of a node. A benefit of BFS on a graph is the ability to determine the smallest cyclic structure containing a given node. In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox-sulfur as the starting, or root node or the search, will discover cyclic bonding structures in order of size. Here we are only concerned with the smallest cyclic structure involving those waters in the first and second hydration shells around the \suldiox. Furthermore, it is possible to reconstruct a cycle's structure by finding its size (number of contributing atoms), and the number of unique waters in the cycle. This allows us to distinguish between various types of cyclic structures encountered.

>>>>

In this work we adopt a nomenclature scheme for \suldiox~in order to quantify hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions from the \suldiox-sulfur to water-oxygens. Thus, an ``SOO'' coordinated \suldiox~molecule forms a single interaction through the sulfur atom to a neighboring water-oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox-oxygens. Analysis of the distribution of \suldiox~coordinations will give insight to how \suldiox~binds to the water surface.

To determine \suldiox~bonding coordinations, we defined intermolecular bonds using the distance criteria of Baer et al.\cite{Baer2010} The bond-length definition is based on a set of distance criteria where a bonding interaction between a \wat-oxygen and \suldiox-sulfur is formed at a distance less than 3.5 \angs, and an \suldiox-oxygen hydrogen bond to a \wat-hydrogen is formed at a distance less than 2.2 \angs.

@@

Hydrated \suldiox~clusters have been studied extensively with many recent experiments and computations forming a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayashi1985, Moin2011, Eckl2008} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The first attempt to elucidate the structure of surface hydrated \suldiox~by Baer et al. showed that the bonding coordination distributions of \suldiox~are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''. In the same work they then focused on the dominant coordination to determine the most likely cluster geometry using two and three-water clusters bound to the \suldiox. However, none of the studies have yet focused directly on the presence of an extended hydrating structure involving closed rings of molecular interactions involving the solvated \suldiox~molecules.

The cluster geometries found in this study imply a cyclic bonding structure through the two or three waters involved. ``Cyclic'' here is used to denote a closed loop formed by the intermolecular hydrogen bonds, S-O interactions, and covalent bonds of the molecules involved. Figure \ref{fig:cyclic-example} depicts one such cyclic structure showing the bonds beginning on the sulfur and returning through a \suldiox-oxygen.

@@

\subsection {Graph Theoretical Details}
The optimized geometry of the \suldiox-hydrates suggests cyclic bonding structures, but it remains a different story entirely when \suldiox~is placed in a dynamic environment such as in the course of MD simulations of an aqueous surface. Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters in a static vacuum. Do the cyclic structures also form in the course of a dynamic bonding process on a simulated water surface, where extended hydrating structures influence \suldiox~and water behavior? To study the formation and behavior of cyclic hydrate structures we employ graph theoretical techniques on MD trajectory data. Previous use of graphs in molecular computations were applied to finding stable arrangements of water clusters, ice, hydrogen bonding, extracting topological molecular properties, and cyclic structure studies.\cite{Anick2002, Huber2007, Radhakrishnan1991, Shi2005, Garcia2004, McDonald1998}

@@

To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A BFS is a recursive algorithm of queuing nodes and all neighboring nodes while performing a specified procedure on each visited node. This is easily performed on adjacency list or connectivity matrix data structures, iterating through nodes (i.e. atoms) of interest in the graph as starting points of the search. In BFS terminology, all nodes are colored during graph traversal to distinguish unvisited nodes (white), queued nodes (gray), and visited nodes (black). Using a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queuing adjacent neighbors of a node. A benefit of BFS on a graph is the ability to determine the smallest cyclic structure containing a given node. In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox-sulfur as the starting, or root node for the search, will discover cyclic bonding structures in order of size. Here we are only concerned with the smallest cyclic structure involving those waters in the first and second hydration shells around the \suldiox. Furthermore, it is possible to reconstruct a cycle's structure by finding its size (number of contributing atoms), and the number of unique waters in the cycle. This allows us to distinguish between various types of cyclic structures encountered.

### 2011-09-08 ComputationalMethods.tex
More of Fred's final edits

On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package, which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).~\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions. The electronic charge density was described using an auxiliary basis set of plane waves. Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP)~\cite{LEE1988}, and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.~\cite{Goedecker1996} A simulation timestep of 1 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the ``cold'' and ``hot'' simulations, respectively. These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.

Initially, 10 equilibrated boxes of side-lengths 10.0\angs, with 36 randomly packed water molecules were used. Five of the boxes were used for each of the cold and hot simulations. A sulfur dioxide molecule was randomly placed onto the surface within 2.5\angs~of a water molecule centrally located above the waters in the z-axis. A copy of the initial system cubes were then expanded along one axis (z-axis) to 25\angs. The system energy was minimized through a geometry optimization. Subsequently, the system was equilibrated for 1 ns in canonical ensemble (NVT) conditions. Periodic boundaries were set on the two short axes to form an infinite slab. The equilibrated systems were then simulated for a further 20 ps in the microcanonical ensemble (NVE), with trajectory snapshots recorded every 1 fs. The initial 1 ns equilibration trajectory was not included in the final analysis. This resulted in 20,000 time steps of system trajectory for analysis in each of the hot and cold replicas of the system.

>>>>

On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package, which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions. The electronic charge density was described using an auxiliary basis set of plane waves. Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP),\cite{LEE1988} and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.\cite{Goedecker1996} A simulation timestep of 1 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the ``cold'' and ``hot'' simulations, respectively. These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.

Initially, 10 equilibrated boxes of side-lengths 10.0\angs, with 36 randomly packed water molecules were used. Five of the boxes were used for each of the cold and hot simulations. A sulfur dioxide molecule was randomly placed onto the surface within 2.5\angs~of a water molecule centrally located above the waters in the z-axis. A copy of the initial system cubes were then expanded along one axis (z-axis) to 25\angs. The system energy was minimized through a geometry optimization. Subsequently, the system was equilibrated for 1 ns in canonical ensemble (NVT) conditions. Periodic boundaries were set on the two short axes to form an infinite slab. The equilibrated systems were then simulated for a further 20 ps in the microcanonical ensemble (NVE), with trajectory snapshots recorded every 1 fs. The initial 1 ns equilibration trajectory was not included in the final analysis. This simulation process resulted in 20,000 time steps of system trajectory for analysis in each of the hot and cold replicas of the system, for a total of 100,000 timesteps at each temperature.
